\documentstyle{article}
\begin{document}

\section*{Multiscale Transforms: A Short Reading List}

\section{Multiscale Transforms}
\begin{itemize}

\item
M.~Antonini, M.~Barlaud, P.~Mathieu, and I.~Daubechies.
\newblock Image coding using wavelet transform.
\newblock {\em IEEE Transactions on Image Processing}, 1(2):205--220, 1992.

\item
J.C. Feauveau.
\newblock {\em Analyse multir\'esolution par ondelettes non-orthogonales et
  bancs de filtres num\'eriques}.
\newblock PhD thesis, Universit\'e Paris Sud, 1990.

\item S.~G. Mallat.
\newblock A theory for multiresolution signal decomposition: the wavelet
  representation.
\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence},
  11(7):674--693, 1989.

\item W.~Sweldens.
\newblock The lifting scheme: {A} construction of second generation wavelets.
\newblock Technical Report 1995:6, Department of Mathematics, University of
  South Carolina, 1995. 


\item M.J. Shensa.
\newblock Discrete wavelet transforms: wedding the \`a trous and Mallat
  algorithms.
\newblock {\em IEEE Transactions on Signal Processing}, 1992.

\end{itemize}


\section{Noise modeling}
\begin{itemize}
 
\item E.~Slezak, F.~Durret, and D.~Gerbal.
\newblock A wavelet analysis search for substructures in eleven X-ray clusters
  of galaxied.
\newblock {\em Astrophysical Journal}, 108(6):1996--2008, 1994.

\item P.~Bury.
\newblock {\em De la distribution de mati\`ere \`a grande \'echelle \`a partir
  des amas d'Abell}.
\newblock PhD thesis, Universit\'e de Nice Sophia Antipolis, Janvier 1995.

\item F.~Murtagh and J.L. Starck.
\newblock Pattern clustering based on noise modeling in wavelet space.
\newblock {\em Pattern Recognition}, 31(7):847--855, 1998.

\item F.~Murtagh and J.L. Starck.
\newblock Image processing through multiscale analysis and measurement noise
  modeling.
\newblock {\em Statistics \& Computing}, 1999.
\newblock to appear.

\item J.L. Starck and F.~Murtagh.
\newblock Automatic noise estimation from the multiresolution support.
\newblock {\em Publications of the Astronomical Society of the Pacific},
  110(744):193--199, 1998.
 
\item J.L. Starck and M.~Pierre.
\newblock Structure detection in low intensity X-ray images.
\newblock {\em Astronomy and Astrophysics, Suppl. Ser.}, 128, 1998.

\item F.~Murtagh and J.L. Starck.
\newblock Overcoming the curse of dimensionality in clustering by means of the
  wavelet transform.
\newblock {\em Computer Journal}, 1999,
\newblock submitted.

\end{itemize}


\section{Compression}
\begin{itemize}
\item  A.~Bijaoui, Y.~Bobichon, and L.~Huang.
\newblock Digital image compression in astronomy -- morphology or wavelets.
\newblock {\em Vistas in Astronomy}, 40(4):587--594, 1996.

\item L.~Huang and A.~Bijaoui.
\newblock Astronomical image data compression by morphological skeleton
  transformation.
\newblock {\em Experimental Astronomy}, 1:311--327, 1991.
 
\item F.~Murtagh, J.L. Starck, and M.~Louys.
\newblock Very high quality image compression based on noise modeling.
\newblock {\em International Journal of Imaging Systems and Technology},
  9:38--45, 1998.

\item R.~White, M.~Postman, and M.~Lattanzi.
\newblock In H.T. MacGillivray and E.B. Thompson, editors, {\em Digitized
  Optical Sky Surveys}, pages 167--175. Kluwer, 1992.

\item J.L. Starck, F.~Murtagh, B.~Pirenne, and M.~Albrecht.
\newblock Astronomical image compression based on noise suppression.
\newblock {\em Publications of the Astronomical Society of the Pacific},
  108:446--455, 1996.

\item M.~Louys, J.-L. Starck, S.~Mei, F.~Bonnarel, and F.~Murtagh.
\newblock Astronomical image compression.
\newblock {\em Astronomy and Astrophysics}, 1999,
\newblock to appear.

\item Y.~Bobichon and A.~Bijaoui.
\newblock A regularized image restoration algorithm for lossy compression in
  astronomy.
\newblock {\em Experimental Astronomy.}, 7(3):239--255, 1997.

\end{itemize}


\section{Detection}
\begin{itemize}
\item F.~Ru\'e and A.~Bijaoui.
\newblock A multiscale vision model to analyse field astronomical images.
\newblock {\em Experimental Astronomy}, 7(3):129--160, 1997.

\item A.~Bijaoui and F.~Ru\'e.
\newblock A multiscale vision model adapted to astronomical images.
\newblock {\em Signal Processing}, 46:345--362, 1995.

\item J.L. Starck, R.~Siebenmorgen, and R.~Gredel.
\newblock Spectral analysis by the wavelet transform.
\newblock {\em Astrophysical Journal}, 482:1011--1020, 1997.

\end{itemize}


\section{Filtering}
\begin{itemize}
\item  A.~Bijaoui, J.L. Starck, and F.~Murtagh.
\newblock Restauration des images multi-echelles par l'algorithme \`a trous.
\newblock {\em Traitement du Signal}, 3:11, 1994.

\item D.L. Donoho and I.M. Johnstone.
\newblock Ideal spatial adaptation by wavelet shrinkage.
\newblock Technical Report 400, Stanford University, 1993.

\item F.~Murtagh, J.L. Starck, and A.~Bijaoui.
\newblock Image restoration with noise suppression using a multiresolution
  support.
\newblock {\em Astronomy and Astrophysics, Suppl. Ser.}, 112:179--189, 1995.

\item F.~Murtagh, J.L. Starck, and A.~Bijaoui.
\newblock Multiresolution in astronomical image processing: A general
  framework.
\newblock {\em The International Journal of Image Systems and Technology},
  6:332--338, 1995.

\item A.~Bijaoui, Y.~Bobichon, Y.~Fang, and F.~Ru\'e.
\newblock M\'ethodes multi\'echelles appliqu\'ees \`a l'analyse des images
  radar \`a ouverture synth\'etique.
\newblock {\em  Traitement du Signal}, 14(2):179--193, 1997.

\item J.L. Starck and F.~Murtagh.
\newblock Image restoration with noise suppression using the wavelet transform.
\newblock {\em Astronomy and Astrophysics}, 288:343--348, 1994.

\item J.L. Starck and F.~Murtagh.
\newblock Multiscale entropy filtering.
\newblock {\em Signal Processing}, 1998,
\newblock to appear.

\item J.L. Starck, F.~Murtagh, and R.~Gastaud.
\newblock A new entropy measure based on the wavelet transform and noise
  modeling.
\newblock {\em \textnormal{Special Issue on Multirate Systems, Filter Banks,
  Wavelets, and Applications of} IEEE Transactions on CAS II}, 45(8), 1998.

\item J.L. Starck and F.~Murtagh.
\newblock Image filtering from multiple vision model combination.
\newblock {\em Image and Vision Computing}, 1999,
\newblock submitted.

\end{itemize}

\section{Deconvolution}

\begin{itemize}
\item  J.P. Burg.
\newblock Annual Meeting International Society Exploratory Geophysics,
  Reprinted in {\em Modern Spectral Analysis}, 
1978, D.G. Childers, ed., IEEE Press,
  New York, 34--41, 1978, 1967.

\item B.R. Frieden.
\newblock {\em {Digital Image Processing}}.
\newblock John Wiley \& Sons, Berlin, Springer-Verlag, 1978.

\item S.F. Gull and J.~Skilling.
\newblock {\em MEMSYS5 User's Manual}, 1991.

\item J.L. Starck and A.~Bijaoui.
\newblock Filtering and deconvolution by the wavelet transform.
\newblock {\em Signal Processing}, 35:195--211, 1994.

\item J.L. Starck and A.~Bijaoui.
\newblock Multiresolution deconvolution.
\newblock {\em Journal of the Optical Society of America A}, 11(4), 1994.
  
\item J.L. Starck, A.~Bijaoui, B.~Lopez, and C.~Perrier.
\newblock Image reconstruction by the wavelet transform applied to aperture
  synthesis.
\newblock {\em Astronomy and Astrophysics}, 283:349--360, 1994.

\item J.L. Starck, A.~Bijaoui, and F.~Murtagh.
\newblock Multiresolution support applied to image filtering and deconvolution.
\newblock {\em CVGIP: Graphical Models and Image Processing}, 57:420--431,
  1995.

\item E.~Pantin and J.L. Starck.
\newblock Deconvolution of astronomical images using the multiscale maximum
  entropy method.
\newblock {\em Astronomy and Astrophysics, Suppl. Ser.}, 315:575--585, 1996.

\end{itemize}


\section{Image registration}

\begin{itemize}
\item J.P. Djamdji, A.~Bijaoui, and R.~Maniere.
\newblock {Geometrical registration of images: the multiresolution approach}.
\newblock {\em Photogrammetric Engineering and Remote Sensing.},
  59(5):645--653, May 1993.

\item  J.P. Djamdji, A.~Bijaoui, and R.~Maniere.
\newblock {Geometrical registration of remotely sensed images with the use of
  the wavelet transform}.
\newblock {\em {SPIE International Symposium on Optical Engineering and
  Photonics}}, pp.\ 412--422, {12-16 April - Orlando} 1993.
\newblock {Volume 1938}.

\item J.P. Djamdji.
\newblock {\em {Analyse en ondelettes et mise en correspondance en
  t\'el\'ed\'etection}}.
\newblock PhD thesis, {Universit\'e de Nice Sophia Antipolis}, {2 D\'ecembre}
  1993.
\end{itemize}

 \section{Wavelet and Fractal analysis}
\begin{itemize}
\item A.~Arneodo, F.~Argoul, E.~Bacry, J.~Elezgaray, and J.~F. Muzy.
\newblock {\em Ondelettes, Multifractales et Turbulences}.
\newblock Diderot, Arts et Sciences, Paris, New-York, Amsterdam, 1995.
\end{itemize}

\end{document}
